---
title: "5.2_strings_and_factors"
author: "Wenshan Qu (wq2160)"
date: "10/21/2021"
output: html_document
---

```{r}
library(tidyverse)
library(rvest)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## String vectors

We create a string vector.
```{r}
string_vec = c("my", "name", "is", "jeff")
```
Find cases where the match exists
```{r}
str_detect(string_vec, "m")
str_detect(string_vec, "jeff")

str_replace(string_vec, "j", "J")
```

Another string vector
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")
str_detect(string_vec, "^i think") ## find "i think" at the beginning of the line
str_detect(string_vec, "i think$") ## find "i think" at the end of the line
```

Another case:
```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Bb]ush") ## find both "Bush" and "bush"
```

Another case:
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]") ## provide a range of letters or numbers that count as a match.
```

Another:
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11") ## you want to search 7 and 11 no matter what's between them!
```

Another:
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[") ## Looking for a special character.
```

Some characters are “special”. These include `[` and `]`, `(` and `)`, and `.`. If you want to search for these, you have to indicate they’re special using `\`. Unfortunately, `\` is also special. 


## Why factors are weird

```{r}
factor_vec = factor(c("male", "male", "female", "female"))
factor_vec

as.numeric(factor_vec)

## If you manually relevel the factor...
factor_vec = fct_relevel(factor_vec, "male")

as.numeric(factor_vec)
```

##NSDUH
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>%
  slice(-1)

table_marj

```

Clean the data

```{r}
data_marj = 
  table_marj %>%
  select(-contains("P Value")) %>%
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
  ) %>% 
  separate(age_year, into = c("age", "year"), "\\(") %>%  ## for sep, we could also `-11`
  mutate(
    year = str_replace(year, "\\)", ""), ## delete the remaining ")"
    percent = str_replace(percent, "[a-c]$", ""), ## clean the percent data values
    percent = as.numeric(percent)
  ) %>% 
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West"))) ## `!` is a complement usage. "except"
```

```{r}
data_marj %>%
  filter(age == "12-17") %>% 
  mutate(
    State = fct_reorder(State, percent)  ## This is not necessary, but it will make the graph more clear!
  ) %>% 
  ggplot(aes(x = State, y = percent, color = year)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90))
```

## Restaurant Inspections

```{r}
data("rest_inspec")
```

`rest_inspec %>%  slice(1:100) %>% view()` helps us to see the first 100 rows of this huge dataset!

```{r}
rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(count = n())

rest_inspec %>% 
  janitor::tabyl(boro, grade)

## A simple filter
rest_inspec = 
rest_inspec %>% 
  filter(
    str_detect(grade, "[ABC]"),
    !(boro == "Missing")
    ) %>% 
  mutate(boro = str_to_title(boro))
```

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  janitor::tabyl(boro, grade)
```

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "The City" = "Manhattan")) %>%
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```

